---
title: "CDW DB Network Plots"
description: |
  Visualizing the VA Corporate Data Warehouse through 2d and 3d network diagrams
date: "`r Sys.Date()`"
author:
  - name: "Richard Evans"
    url: https://github.com/rrevansUM
    affiliation: Center for Clinical Management Research, U.S. Dept. of Veterans Affairs
    affiliation_url: https://www.annarbor.hsrd.research.va.gov/
output: 
  distill::distill_article:
    toc: true
---

```{r setup, include = FALSE}
library(knitr)

options(max.print = "200")
opts_chunk$set(echo    = TRUE,
	             cache   = FALSE,
               prompt  = FALSE,
               tidy    = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE)
opts_knit$set(width = 80)
```

# Introduction

Network plots depicting relationships between Database tables and views in CDW, the corporate data warehouse. There are 2 sources of meta-data, that which I have created (CDW), and that which can be downloaded from `CDWWork.Meta.DWViewForeignKey`^[see SQL code in appendix], both are incomplete to some extent. I will work with just the one queried from CDW directly, as it's more complete.

# Import Data

Relationships (**edges**) and tables (**nodes**) imported from SQL DB

<font size="4">Edges</font>

```{r}
library(readxl)
library(tidyverse)
library(magrittr)

file <- "Data/CDW_Relationships.xlsx" # nodes and edges in separate sheets

edges <- read_xlsx(file, sheet = "metaVINCI")
```

We're going to create nodes from the edges

# Process Data for Network Setup

```{r}
nodes1 <- edges %>% distinct(Entity1) %>% rename(id = Entity1)
nodes2 <- edges %>% distinct(Entity2) %>% rename(id = Entity2)
nodes <- bind_rows(nodes1, nodes2)
nodes$label <- nodes$id
nodes$Domain <- sub("_.*", "", nodes$label)
nodes <- nodes %>%
  mutate(
    Domain = case_when(
      Domain == "ADR"          ~ "Active Directory",
      Domain == "Appt"         ~ "Appointment",
      Domain == "Chem"         ~ "LabChem",
      Domain == "Con"          ~ "Consult",
      Domain == "Fee"          ~ "Purchased Care",
      Domain == "HF"           ~ "Health Factor",
      Domain == "IB"           ~ "Integrated Billing",
      Domain == "Immun"        ~ "Immunization",
      Domain == "Inpat"        ~ "Inpatient",
      Domain == "MH"           ~ "Mental Health",
      Domain == "Micro"        ~ "Microbiology",
      Domain == "NonVAMed"     ~ "Non-VA Medication",
      Domain == "Outpat"       ~ "Outpatient",
      Domain == "Patsub"       ~ "Patient",
      Domain == "Rad"          ~ "Radiology",
      Domain == "SPatientText" ~ "SPatient",
      Domain == "StaffSub"     ~ "Staff",
      Domain == "Surg"         ~ "Surg",
      Domain == "Vital"        ~ "Vitals",
      Domain == "WH"           ~ "Womens Health",
      TRUE ~ Domain
    ),
    group = ifelse(Domain == "Dim" | Domain == "NDim", "Dim", "Fact")
  ) %>% 
  distinct()

# Edges need renaming
edges <- edges %>% rename(from = Entity1, to = Entity2)
```


# Network Visualization

## 2-Dimensional

```{r}
library(visNetwork)

test.edges <- edges %>% filter(Domain == "Dim")

nodes.in.net <- data.frame(
  id = c(unique(test.edges$from), unique(test.edges$to))
)

test.nodes <- nodes %>%
  right_join(nodes.in.net, by = "id") %>%
  select(-Domain) %>%
  distinct()

visNetwork(test.nodes, test.edges) %>%
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = TRUE
  ) %>%
  visInteraction(navigationButtons = TRUE)
```


# SQL Code {.appendix}

```{r, echo = TRUE, eval = FALSE}

/******************************* Relationships *******************************/

USE ORD_Sears_201803016D
GO

SELECT COUNT(*) AS MetaDWViewForeignKey
	FROM CDWWork.Meta.DWViewForeignKey
-- 2686 relationships in total

SELECT a.TABLE_NAME
	INTO #SearsViews
		FROM INFORMATION_SCHEMA.COLUMNS AS a
			 JOIN INFORMATION_SCHEMA.TABLES AS t 
				ON a.TABLE_NAME = t.TABLE_NAME
				   AND 
				   a.TABLE_SCHEMA = t.TABLE_SCHEMA
		WHERE TABLE_TYPE = 'view'
		GROUP BY a.TABLE_NAME
-- nrow = 237, ncol = 1

-- create temporary version of Meta.DWViewForeignKey, concatenating table names
SELECT FKSchemaName    + '_' + FKViewName AS FKFullName,
	     PKSchemaName    + '_' + PKViewName AS PKFullName,
	     FKViewFieldName + ' = ' + PKViewFieldName AS Link
	INTO #Relationships
		FROM CDWWork.Meta.DWViewForeignKey
-- nrow = 2686, ncol = 3

-- left join to subset to Sears cohort views
SELECT TABLE_NAME AS Entity1,
       PKFullName AS Entity2,
	     Link
	 INTO gen.Relationships
		  FROM #SearsViews AS a
			    LEFT JOIN #Relationships AS b ON a.TABLE_NAME = b.FKFullName
-- nrow = 1104, ncol = 3

-- clean temp table workspace

SELECT LEFT(name, charindex('_', name) - 1) AS TempTables
	FROM tempdb..sysobjects
	WHERE charindex('_', name) > 0 
		  AND xtype = 'u' 
		  AND NOT object_id('tempdb..' + name) IS NULL
	ORDER BY TempTables

DROP TABLE #Relationships, #SearsViews

/****************************** End of Document ******************************/

```

# R-Session Info {.appendix}

```{r, echo = FALSE, eval = TRUE}
sessionInfo()
```

